Size-differentiated Export Flux in Different Dynamical Regimes in the Ocean

### 2020-12-04 size_differentiated_export.tex
Updates from Overleaf
final adjusmtents
hopefully.

\documentclass[article,linenumbers]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

@@

\usepackage{amsmath}
\usepackage{ulem}

@@

	Photosynthesis in the sunlit upper ocean and the production of Particulate Organic Carbon (POC) takes up dissolved inorganic carbon and facilitates the uptake of CO$_2$ from the atmosphere. The sinking of POC exports organic carbon from the upper ocean to the interior, leading to the sequestration of carbon \cite{Falkowski_1998} on timescales ranging from days to years depending on the sinking depth and  circulation. Despite progress on sampling and viewing particles in the ocean, direct measurements of particles sinking velocities are difficult to obtain, and often inferred from key parameters such as particle type, size, and density \cite{McDonnell_2010,Mcdonnell_2012}.

@@

	Recent studies have shown that ocean dynamics can play a role in driving the transport of carbon from the euphotic layer to the ocean interior.  For example, enhanced vertical velocities along the edge of a mesoscale eddy led to a funneling of particles along the eddy's periphery \cite{vanHaren_2006, Waite_2016} and in mesoscale features in the California current \cite{Stukel_2017}. \citeA{Omand_2015} found that submesoscale mixed layer eddies, while contributing to the restratification of a frontal zone, were subducting a large amount of non-sinking POC from the surface productive layer during the onset of the Spring bloom in the subpolar North Atlantic. Advectively subducting plumes or filaments of high oxygen, chlorophyll and small POC (evidenced through backscatter) were detected from a suite of gliders during the North Atlantic Bloom experiment \cite{Alkire_2012}. Using model simulations to capture the process of eddy-driven subduction, \citeA{Omand_2015} estimated the downward advective flux of non-sinking POC and parameterized it. \citeA{Briggs_2011} quantified the flux of fast-sinking particles consisting largely of diatoms from observations of optical backscatter. But, these estimates did not account for a range of sinking particle velocities.  Typically, POM has a wide spectrum of sinking velocities and in order to understand its fate and export, we need to consider the biomass distribution as a function of the particle sinking velocity spectrum and its interaction with the dynamics of the flow field in the ocean.

	A growing body of literature focusing on submesoscale (1-10 km) dynamics is exploring its impact on biogeochemical processes \cite{Levy_2012, Mahadevan_2016}. Submesoscale dynamics, characterized by Rossby numbers of order 1, typically develop in filaments in areas where sharp density fronts exist \cite{Thomas_2013b,Klein_2009,McWilliams_2016}. In this dynamical regime, geostrophic balance breaks down and a secondary ageostrophic circulation develops across the front, capable of generating large vertical velocities on the order of 100 m/day \cite{FoxKemper_2008,Mahadevan_2016}. On the denser side of the front, the vorticity is cyclonic and associated with downwelling velocities, while anticyclonic vorticity and upwelling is expected on the lighter side of the front. The distribution of relative vorticity associated with submesoscale dynamics near the surface exhibits an asymmetry with higher values of positive vorticity than negative vorticity \cite{Rudnick_2001}, leading to more localized and more intense downwelling regions, as opposed to weaker and larger scale upwelling regions \cite{Mahadevan_2006}. Enhanced vertical velocities can aid the supply nutrients to the sunlit layer of the ocean for primary production \cite{Mahadevan_2000,Levy_2001} or can significantly increase the export of POC to the ocean interior through localized downwelling \cite{Levy_2012, Gruber_2011,Estapa_2015,Omand_2015}. The downwelling velocities $\mathcal{O}$(100 m/day) generated at submeso-scales provide a physical mechanism for exporting slow sinking or neutrally buoyant particles on timescales shorter than their remineralization timescales. If the fraction of  biomass associated with such slow sinking velocities is significant, submesoscale dynamics can potentially impact the export of POC.

@@

	Studies investigating submesoscale dynamics traditionally focused on regions where the presence of submesoscale fronts and filaments are established, such as western boundary currents with strong gradients \cite{Dasaro_2011, Thomas_2013}, or the edge of mesoscale features \cite{vanHaren_2006,Waite_2016}. The seasonality in submesoscale dynamics captured in the glider dataset at Station Papa and reflected in the model experiments, echoes the behavior seen from recent observational studies conducted at a similar latitude in the Atlantic Ocean, which demonstrate the intensification of submesoscale dynamics in the wintertime \cite{Thompson_2016, Buckingham_2016}. Despite being sometimes qualified as an ``eddy desert" with low kinetic energy \cite{Chelton_2011}, ocean characteristics in the eastern part of the Pacific subpolar gyre suggest the presence of submesoscale features in the wintertime: strong density gradients, localized Rossby numbers of order 1, a balanced Richardson number $Ri_b = \frac{f^2N^2}{M^4}$ smaller than 1, a positively skewed distribution in vorticity, and a negatively skewed distribution of vertical velocities \cite<see Figure \ref{fig: dynamics};>[]{Thomas_2013b,Rudnick_2001, Buckingham_2016}.

@@

	\section*{Acknowledgments}
	The work was funded by NASA grant NNX16AR48G, to complement the EXport Processes in the global Ocean from RemoTe Sensing (EXPORTS) program. We would like to thank A. Thompson (Caltech) for his useful comments and suggestions, N. Pelland (NOAA) and C. Eriksen (University of Washington) for sharing the glider data collected at Ocean Station Papa, as well as S. Essink for his assistance in developing the particle-tracking code. All data used in this manuscript are publicly available: Ocean Station Papa data is available on PMEL's website \cite<\url{https://www.pmel.noaa.gov/ocs/Papa};>[]{PMEL_data} and gridded Argo products can be downloaded at \url{http://www.seanoe.org/data/00348/45945} \cite{Gaillard_2015}. Glider data is archived at the University of Washington's Library \cite<\url{http://hdl.handle.net/1773/41656};>[]{Pelland_2018_data}. Code to reproduce analysis and figures are publicly available at \url{https://github.com/matdever/Size-differentiated_Export_GBC} \cite{Dever_dataset}. Due to the very large filesize, model outputs and particle trajectories are available on request. %Finally, we are grateful to the two anonymous reviewers for their valuable comments and suggestions.

>>>>

\documentclass[article,linenumbers]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[finalnew]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

@@

\usepackage{amsmath}
%\usepackage{ulem}

@@

	Photosynthesis in the sunlit upper ocean and the production of Particulate Organic Carbon (POC) takes up dissolved inorganic carbon and facilitates the uptake of CO$_2$ from the atmosphere. The sinking of POC exports organic carbon from the upper ocean to the interior, leading to the sequestration of carbon \cite{Falkowski_1998, Boyd_2019} on timescales ranging from days to years depending on the sinking depth and  circulation. Despite progress on sampling and viewing particles in the ocean, direct measurements of particles sinking velocities are difficult to obtain, and often inferred from key parameters such as particle type, size, and density \cite{McDonnell_2010,Mcdonnell_2012}.

@@

	Recent studies have shown that ocean dynamics can play a role in driving the transport of carbon from the euphotic layer to the ocean interior \cite{Stukel_2017b,Stukel_2018,Llort_2018}.  For example, enhanced vertical velocities along the edge of a mesoscale eddy led to a funneling of particles along the eddy's periphery \cite{vanHaren_2006, Waite_2016} and in mesoscale features in the California current \cite{Stukel_2017}. \citeA{Omand_2015} found that submesoscale mixed layer eddies, while contributing to the restratification of a frontal zone, were subducting a large amount of non-sinking POC from the surface productive layer during the onset of the Spring bloom in the subpolar North Atlantic. Advectively subducting plumes or filaments of high oxygen, chlorophyll and small POC (evidenced through backscatter) were detected from a suite of gliders during the North Atlantic Bloom experiment \cite{Alkire_2012}. Using model simulations to capture the process of eddy-driven subduction, \citeA{Omand_2015} estimated the downward advective flux of non-sinking POC and parameterized it. \citeA{Briggs_2011} quantified the flux of fast-sinking particles consisting largely of diatoms from observations of optical backscatter. But, these estimates did not account for a range of sinking particle velocities. Typically, POM has a wide spectrum of sinking velocities and in order to understand its fate and export, we need to consider the biomass distribution as a function of the particle sinking velocity spectrum and its interaction with the dynamics of the flow field in the ocean.

	A growing body of literature focusing on submesoscale (1-10 km) dynamics is exploring its impact on biogeochemical processes \cite{Levy_2012, Mahadevan_2016}. Submesoscale dynamics, characterized by Rossby numbers of order 1, typically develop in filaments in areas where sharp density fronts exist \cite{Thomas_2013b,Klein_2009,McWilliams_2016}. In this dynamical regime, geostrophic balance breaks down and a secondary ageostrophic circulation develops across the front, capable of generating large vertical velocities on the order of 100 m/day \cite{FoxKemper_2008,Mahadevan_2016}. On the denser side of the front, the vorticity is cyclonic and associated with downwelling velocities, while anticyclonic vorticity and upwelling is expected on the lighter side of the front. The distribution of relative vorticity associated with submesoscale dynamics near the surface exhibits an asymmetry with higher values of positive vorticity than negative vorticity \cite{Rudnick_2001}, leading to more localized and more intense downwelling regions, as opposed to weaker and larger scale upwelling regions \cite{Mahadevan_2006}. Enhanced vertical velocities can aid the supply nutrients to the sunlit layer of the ocean for primary production \cite{Mahadevan_2000,Levy_2001} or can significantly increase the export of POC to the ocean interior through localized downwelling \cite{Levy_2012, Gruber_2011,Estapa_2015,Omand_2015}, \add{a process that might not be captured if submesoscales O(1 km) are not fully resolved} \cite{Resplandy_2019}. The downwelling velocities $\mathcal{O}$(100 m/day) generated at submeso-scales provide a physical mechanism for exporting slow sinking or neutrally buoyant particles on timescales shorter than their remineralization timescales. If the fraction of  biomass associated with such slow sinking velocities is significant, submesoscale dynamics can potentially impact the export of POC.

@@

	Studies investigating submesoscale dynamics traditionally focused on regions where the presence of submesoscale fronts and filaments are established, such as western boundary currents with strong gradients \cite{Dasaro_2011, Thomas_2013}, or the edge of mesoscale features \cite{vanHaren_2006,Waite_2016}. The seasonality in submesoscale dynamics captured in the glider dataset at Station Papa and reflected in the model experiments, echoes the behavior seen from recent observational studies conducted at a similar latitude in the Atlantic Ocean, which demonstrate the intensification of submesoscale dynamics in the wintertime \cite{Thompson_2016, Buckingham_2016, Erickson_2018}. Despite being sometimes qualified as an ``eddy desert" with low kinetic energy \cite{Chelton_2011}, ocean characteristics in the eastern part of the Pacific subpolar gyre suggest the presence of submesoscale features in the wintertime: strong density gradients, localized Rossby numbers of order 1, a balanced Richardson number $Ri_b = \frac{f^2N^2}{M^4}$ smaller than 1, a positively skewed distribution in vorticity, and a negatively skewed distribution of vertical velocities \cite<see Figure \ref{fig: dynamics};>[]{Thomas_2013b,Rudnick_2001, Buckingham_2016}.

@@

	\section*{Acknowledgments}
	The work was funded by NASA grant NNX16AR48G, to complement the EXport Processes in the global Ocean from RemoTe Sensing (EXPORTS) program. We would like to thank A. Thompson (Caltech) for his useful comments and suggestions, N. Pelland (NOAA) and C. Eriksen (University of Washington) for sharing the glider data collected at Ocean Station Papa, as well as S. Essink for his assistance in developing the particle-tracking code. All data used in this manuscript are publicly available: Ocean Station Papa data is available on PMEL's website \cite<\url{https://www.pmel.noaa.gov/ocs/Papa};>[]{PMEL_data} and gridded Argo products can be downloaded at \url{http://www.seanoe.org/data/00348/45945} \cite{Gaillard_2015}. Glider data is archived at the University of Washington's Library \cite<\url{http://hdl.handle.net/1773/41656};>[]{Pelland_2018_data}. Both the code and dataset necessary to reproduce analysis and figures are publicly available\cite{Dever_dataset, Dever_dataset2}. Finally, we are grateful to the two anonymous reviewers for their valuable comments and suggestions.
